---
title: "Maps_with_r"
author: "Mahmudur Rahman"
date: "05/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


load libraries
```{r}
library(tidyverse)
library(magrittr)
theme_set(theme_minimal())
```

Geographical Information System (GIS)

Choropleth maps show geographical regions colored, shaded, or graded according to some variable

geocomputation with r Book:[<https://geocompr.robinlovelace.net/>]

spatial data , spacial distribution in a compeling way .
Cartography is the study and practice of making and using maps.


#choropleth maps with ggplot2

## Map U.S. state-level data

```{r}
election %>% select(state, total_vote,
                    r_points, pct_trump, party, census) %>%
    sample_n(size = 5)

election$fips#  FIPS code is a federal code that numbers states and territories of the US.
```

```{r}
# Hex color codes for Dem Blue and Rep Red
party_colors <- c("#2E74C0", "#CB454A")

p0 <- ggplot(data = subset(election, st %nin% "DC"), # %nin% = not in 
             aes(x = r_points,
                           y = reorder(state, r_points),
                           color = party))
p1 <- p0+
geom_vline(xintercept = 0,color= "grey30")+
  geom_point(size =2)
  
p2 <- p1+
  scale_color_manual(values = party_colors)
  
p3 <- p2 +
  scale_x_continuous(breaks = c(-30, -20, -10, 0, 10, 20, 30, 40),
                     labels = c("30\n(Clinton)", "20","10","0","10","20","30", "40\n(Trump)"))
                  
 p3 +
   facet_wrap(~ census,ncol = 1,scales = "free_y")+
   guides(color = F)+
   labs(x = "Margin Point",
        y = "")+
   theme(axis.text = element_text(size = 8))

```

the first task in drawing a map is to get a data frame with the right information in it, and in the right order. First we load R’s maps package, which provides us with some pre-drawn map data

blog with London map[https://towardsdatascience.com/plotting-a-map-of-london-crime-data-using-r-8dcefef1c397]

```{r}
library(maps)
us_states <- map_data("state")
head(us_states)
dim(us_states)


```

We can make a blank state map with this data, using geom_polygon()
```{r}
p <- ggplot(data = us_states,
            mapping = aes(x = long, y = lat,
                          group = group,fill = region))

p1 <- p + geom_polygon( color = "grey90",size = .1)+
  guides(fill =F)

#Improving the projection
p1+
   coord_map(projection = "albers", lat0 = 39, lat1 = 45)

```

data prep
```{r}
election$region <- tolower(election$state)
us_states_elec <- left_join(us_states, election)

us_states_elec %>% 
  select(1:6) %>% 
  sample_n(5)
  
```

plot in map with newly created df
```{r}
p1 <-   us_states_elec %>% 
  ggplot(aes(long,lat,group = group, fill = party))+
  geom_polygon(color = "gray90", size = 0.1) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45)

p2 <- p1+
  scale_fill_manual(values = party_colors)+
  labs(title = "Election Results 2016", fill = NULL)

p2+
  theme_map()# removes grid lines /axis labels
  
  
```







